\documentclass[11pt]{article}

% arXiv submission info
% Primary category: cs.SE (Software Engineering)
% Secondary category: cs.AI (Artificial Intelligence)

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{positioning, shapes.geometric, arrows.meta, calc}

% Custom command for placeholders
\newcommand{\placeholder}[1]{\textcolor{red}{\textbf{[#1]}}}
\newcommand{\todo}[1]{\textcolor{orange}{\textbf{[TODO: #1]}}}
\newcommand{\metric}[1]{\textsc{#1}}

\title{Compiled AI: Deterministic Code Generation for\\LLM-Based Workflow Automation}

\author{
  Geert Trooskens\textsuperscript{1} \quad
  Aaron Karlsberg\textsuperscript{1} \quad
  Anmol Sharma\textsuperscript{1} \quad
  Lamara De Brouwer\textsuperscript{1} \\[0.3em]
  Max Van Puyvelde\textsuperscript{2} \quad
  Walter A. De Brouwer\textsuperscript{2}
  \\[1em]
  \textsuperscript{1}XY.AI Labs, Palo Alto, CA \\
  \textsuperscript{2}Stanford University School of Medicine, Stanford, CA \\[0.5em]
  \texttt{foundation@xy.ai}
}

\date{January 2026}

\begin{document}

\maketitle

% ============================================================================
\begin{abstract}
% ============================================================================

We study \emph{compiled AI}, a paradigm in which large language models generate executable code artifacts during a compilation phase, after which workflows execute deterministically without further model invocation. By constraining generation to narrow business-logic functions embedded in validated templates, compiled AI trades runtime flexibility for predictability, auditability, and cost efficiency, while retaining the ability to improve artifacts through observed execution patterns.

We introduce (i) a system architecture for constrained LLM-based code generation, (ii) a four-stage generation-and-validation pipeline that converts probabilistic model output into production-ready code artifacts, and (iii) an evaluation framework that measures operational metrics including token amortization, determinism, reliability, security, and cost. In preliminary evaluation on function-calling tasks (n=499 across two BFCL categories), compiled AI achieves 96\% task completion with zero execution tokens, breaking even with runtime inference at approximately 21 transactions and reducing token consumption by 47$\times$ at 1,000 transactions. Security evaluation across 135 test cases demonstrates 96.7\% accuracy on prompt injection detection and 87.5\% on static code safety analysis with zero false positives.

\end{abstract}

% ============================================================================
\section{Introduction}
\label{sec:introduction}
% ============================================================================

Large language models (LLMs) are increasingly deployed to automate enterprise workflows, evolving from question-answering systems toward autonomous agent architectures. Recent frameworks enable LLMs to reason over tools, APIs, and multi-step tasks at runtime, often through conversational or multi-agent interaction patterns \citep{autogen2024, crewai2024, langchain2022}. While these approaches demonstrate flexibility, they rely on repeated model invocation during execution, leading to high token consumption, variable latency, and non-deterministic behavior.

Empirical studies and industrial deployments have highlighted persistent reliability challenges in runtime agent systems, particularly in multi-step business workflows where success rates degrade due to specification ambiguity, inter-agent coordination failures, and stochastic model outputs \citep{cemri2025multiagent, atil2024nondeterminism}. These limitations pose obstacles to deployment in enterprise environments, where determinism, auditability, and cost predictability are often strict requirements.

This paper investigates an alternative design point: \emph{compiled AI}, a paradigm in which LLMs generate executable code artifacts during a one-time compilation phase, after which workflows execute deterministically without further model inference. The key observation motivating this work is that many enterprise workflows require intelligence to \emph{design} but not to \emph{execute} repeatedly. Once business logic is specified and validated, repeated execution can be handled by conventional software infrastructure, with full observability that enables targeted refinement of the compiled artifacts over time.

Compiled AI represents a trade-off between flexibility and determinism. Rather than interpreting natural language or reasoning at runtime, an LLM is constrained to generate narrow business-logic functions within pre-validated templates. These generated artifacts are subjected to a multi-stage validation pipeline (security analysis, syntactic verification, execution testing, and output accuracy checks) before deployment. After validation, workflows execute as static code with predictable latency, zero stochasticity, and near-zero marginal inference cost.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[scale=0.9]
    % Axes
    \draw[->] (0,0) -- (10,0) node[right] {\small Transactions ($n$)};
    \draw[->] (0,0) -- (0,6) node[above] {\small Total Tokens};

    % Grid lines (light)
    \foreach \x in {2,4,6,8} {
        \draw[gray!20] (\x,0) -- (\x,5.5);
    }
    \foreach \y in {1,2,3,4,5} {
        \draw[gray!20] (0,\y) -- (9.5,\y);
    }

    % Y-axis labels (actual token values: 0-500K)
    \node[left, font=\scriptsize] at (0,1) {100K};
    \node[left, font=\scriptsize] at (0,2) {200K};
    \node[left, font=\scriptsize] at (0,3) {300K};
    \node[left, font=\scriptsize] at (0,4) {400K};
    \node[left, font=\scriptsize] at (0,5) {500K};

    % Runtime inference line (linear growth: 500 tokens/tx)
    % At n=1000, runtime = 500K tokens = 5 units on y-axis
    \draw[thick, red!70] (0,0) -- (9.5,4.75);
    \node[red!70, right] at (9.5, 4.75) {\small Runtime (500/tx)};

    % Compiled AI line (flat at 11K tokens = 0.11 units on y-axis)
    \draw[thick, blue!70] (0,0.11) -- (9.5,0.11);
    \node[blue!70, right] at (9.5, 0.11) {\small Compiled (11K)};

    % Generation cost annotation
    \draw[<->, gray] (-0.5,0) -- (-0.5,0.11);
    \node[gray, left, font=\scriptsize, align=right] at (-0.5,0.055) {11K\\gen cost};

    % Break-even point at n*=21 (x=0.20 on scale where 9.5=1000)
    % x position: 21/1000 * 9.5 = 0.20
    \draw[dashed, gray] (0.20,0) -- (0.20,0.11);
    \node[below, font=\scriptsize] at (0.20,-0.1) {$n^*\!=\!21$};
    \fill[black] (0.20,0.11) circle (2pt);
    \node[above, font=\scriptsize] at (0.5,0.3) {Break-even};

    % Savings region
    \fill[green!10] (0.20,0.11) -- (9.5,0.11) -- (9.5,4.75) -- (0.20,0.11);
    \node[green!50!black, font=\scriptsize] at (6,2.2) {47$\times$ savings};
    \node[green!50!black, font=\scriptsize] at (6,1.8) {at $n$=1000};

    % X-axis labels (actual transaction counts)
    \node[below, font=\scriptsize] at (0.95,0) {100};
    \node[below, font=\scriptsize] at (4.75,0) {500};
    \node[below, font=\scriptsize] at (9.5,0) {1000};

    % Data points for runtime (actual measurements)
    \fill[red!70] (0.095,0.0475) circle (1.5pt); % n=10
    \fill[red!70] (0.19,0.095) circle (1.5pt); % n=20
    \fill[red!70] (0.95,0.475) circle (1.5pt); % n=100
    \fill[red!70] (9.5,4.75) circle (1.5pt); % n=1000

\end{tikzpicture}
\caption{Token consumption comparison: runtime inference vs. compiled AI (preliminary data from BFCL function-calling benchmark, n=499). Runtime inference consumes 500 tokens per transaction. Compiled AI incurs a one-time $\sim$11,000 token generation cost, then executes deterministically at zero marginal token cost. Break-even occurs at $n^*\approx21$ transactions; at 1,000 transactions, compiled AI reduces token consumption by 47$\times$.}
\label{fig:economics}
\end{figure}

\paragraph{Contributions.}
This paper makes three contributions:
\begin{itemize}
    \item We introduce an architecture for constrained LLM-based code generation in which models produce bounded business-logic functions embedded in durable workflow templates, enabling deterministic execution (Section~\ref{sec:architecture}).
    \item We present a four-stage generation-and-validation pipeline that empirically converts probabilistic model output into production-ready code artifacts through iterative regeneration (Section~\ref{sec:architecture}).
    \item We define an evaluation framework that measures the operational viability of LLM-based workflow systems using metrics for token amortization, determinism, reliability, validation effectiveness, and cost (Section~\ref{sec:framework}).
\end{itemize}

Our goal is not to replace runtime inference in all settings, but to provide a principled basis for identifying regimes (such as well-specified, high-volume workflows) where compilation yields operational advantages.

% ============================================================================
\section{Related Work}
\label{sec:related-work}
% ============================================================================

\paragraph{Novelty and Scope.}
Prior work on LLM-based agents, program-aided reasoning, and workflow synthesis has focused primarily on model capability: whether systems can generate or reason over programs. In contrast, this work treats production viability as a first-class concern. Our contributions are not a new code generation algorithm per se, but a systems-oriented study of how probabilistic LLM outputs can be converted into deterministic, auditable execution artifacts. To our knowledge, this is the first systematic evaluation of LLM-based workflow automation using operational and economic metrics such as token amortization, execution determinism, validation convergence, and cost at scale.

\subsection{The Reliability Crisis in LLM Agents}

Recent empirical studies reveal a fundamental reliability problem in multi-agent LLM systems. \citet{cemri2025multiagent} analyzed 1,600+ annotated traces across seven major frameworks (MetaGPT, ChatDev, AutoGen), finding that \textbf{79\% of failures stem from specification and coordination issues}, not infrastructure limitations. Their Multi-Agent System Failure Taxonomy (MAST) identifies specification failures (41.8\%) and inter-agent coordination failures (36.9\%) as dominant failure modes, with infrastructure problems accounting for only 16\% of breakdowns.

Production deployments confirm this pattern. Salesforce's CRMArena-Pro benchmark shows success rates degrading from 58\% (single-turn) to \textbf{35\% in multi-turn interactions} \citep{salesforce2025agents}. \citet{pan2025measuring} surveyed 306 practitioners across 26 domains and found that 68\% limit agents to $\leq$10 autonomous steps before requiring human intervention, with 74\% depending primarily on human evaluation rather than automated metrics. Reliability was cited as the primary deployment challenge.

Critically, non-determinism persists even with temperature=0. \citet{ouyang2023nondeterminism} demonstrated 18-75\% output variance in code generation tasks due to architectural factors including Mixture-of-Experts routing and floating-point variance, fundamentally undermining reproducibility in interpreted AI approaches. \citet{atil2024nondeterminism} found accuracy varies up to 15\% across runs at temperature=0, with the gap between best and worst possible performance reaching 70\%.

These findings motivate our core architectural choice: removing the LLM from the execution loop entirely, confining model invocation to a one-time compilation phase where variance can be detected and corrected before deployment.

\subsection{Benchmark Saturation and Contamination}

The apparent progress on code generation benchmarks masks significant methodological concerns. HumanEval is now saturated ($>$95\% for frontier models), with Stanford's AI Index 2025 formally retiring it for frontier evaluation. More troubling, contamination studies reveal 8-18\% of HumanEval appearing in pre-training data \citep{yang2023rethinking}, with average 39.4\% performance drops after decontamination \citep{xu2024contamination}.

\begin{table}[ht]
\centering
\caption{Code generation benchmark performance, December 2025. HumanEval is saturated; contamination-resistant benchmarks reveal substantial capability gaps.}
\label{tab:benchmarks}
\begin{tabular}{llccc}
\toprule
Benchmark & Model & Score & Clean Score & Date \\
\midrule
\multirow{2}{*}{SWE-Bench Verified} & Claude Opus 4.5 & 80.9\% & --- & Nov 2025 \\
 & GPT-5.2 & 80.0\% & --- & Dec 2025 \\
\midrule
SWE-Bench Pro & Claude Opus 4.1 & --- & 22.7\% & Jan 2026 \\
 & GPT-5 & --- & 23.3\% & Jan 2026 \\
\midrule
HumanEval & Claude Opus 4.5 & 96.4\% & --- & Nov 2025 \\
\midrule
BigCodeBench & GPT-4o & 60.0\% & --- & 2024 \\
\bottomrule
\end{tabular}
\end{table}

``The SWE-Bench Illusion'' \citep{shi2025swebenchillusion} provides striking evidence that models achieve 76\% accuracy on path identification from issue text alone, with performance dropping to \textbf{53\% on repositories outside SWE-Bench}, suggesting memorization rather than generalization. When SWE-Bench Pro introduced contamination-resistant benchmarks with GPL-licensed repositories, top model performance collapsed from 70\%+ to \textbf{$\sim$23\%}.

The synthetic-to-real gap is severe: models achieving 84-89\% on synthetic benchmarks attain only \textbf{25-34\% on real-world class-level tasks} \citep{rao2025synthetic}. Perhaps most striking, the METR developer productivity study found that developers \emph{perceived} 20\% productivity gains while objective measurements showed \textbf{19\% slower} task completion \citep{metr2025productivity}, directly contradicting adoption narratives.

These findings suggest that capability benchmarks alone are insufficient for evaluating production systems. Our evaluation framework (Section~\ref{sec:framework}) therefore emphasizes operational metrics (token amortization, execution determinism, and validation convergence) over generation accuracy alone.

\subsection{Constrained Generation and Formal Verification}

Structured generation techniques offer a path toward reliable outputs. SynCode \citep{ugare2024syncode} achieves \textbf{96\% reduction in syntax errors} through grammar-constrained decoding using offline-constructed DFA mask stores. Type-constrained decoding \citep{mundler2025typeconstrained} reduces compilation errors by $>$50\% while improving functional correctness. XGrammar \citep{dong2024xgrammar} demonstrates production-viable performance with up to 100$\times$ speedup over existing constrained decoding approaches.

Formal verification pipelines show promise for critical systems. Astrogator \citep{councilman2025astrogator} verifies correct code in 83\% of cases and identifies incorrect code in 92\% using a Formal Query Language for intent specification. The Guaranteed Safe AI framework \citep{dalrymple2024guaranteed} provides theoretical foundations for verification without requiring interpretability by combining world models, safety specifications, and formal verifiers.

However, fundamental limitations remain: \citet{xu2024hallucination} demonstrate that structural hallucinations cannot be eliminated by larger training sets alone, positioning constrained decoding as complementary rather than complete. Our approach combines constrained generation with multi-stage validation, treating the pipeline as a whole rather than relying on any single technique.

For healthcare specifically, \citet{neupane2025hipaa} present a framework for HIPAA-compliant agentic AI systems, integrating attribute-based access control with PHI sanitization pipelines, an approach our template-based architecture can encode as compliance-by-construction.

\subsection{The Bitter Lesson and Its Limits}

Sutton's ``Bitter Lesson'' \citep{sutton2019bitter} argues that general methods leveraging computation ultimately win over hand-crafted approaches. Code Llama's results support this: a 7B code-specialized model outperforms Llama 2 70B on HumanEval/MBPP \citep{roziere2023codellama}. Karpathy's ``Software 2.0'' thesis \citep{karpathy2017software2} characterized neural networks as an alternative programming paradigm. More recently, Karpathy observed that leading practitioners are converging on ``spec-driven development'' (writing specifications while AI handles implementation), which he characterizes as ``the limit of imperative to declarative transition'' \citep{karpathy2026specdriven}. This trajectory aligns directly with compiled AI: specifications in, deterministic code out.

However, recent work questions whether scale alone is sufficient. The critical constraint may be data, not compute: scaling laws show $C \sim D^2$ (compute scales quadratically with data), but high-quality training data is finite (approximately 10T curated tokens), ``and there is no second Internet'' \citep{chakrabarti2025bitter}. Purpose-built hybrid systems can outperform pure neural approaches: Stockfish demonstrates this in chess, and HRM (27M parameters) beats o3-mini-high on ARC-AGI-1 \citep{breunig2025bitter}.

Our work operationalizes a hybrid insight: LLMs excel at \emph{understanding intent} but struggle with \emph{reliable execution}. By treating the LLM as a compiler and front-loading computation to the generation phase, we leverage model capability for design while relying on conventional software for deterministic execution.

\subsection{The Compilation Paradigm}

The most directly relevant work validates compilation as a distinct paradigm. DSPy \citep{khattab2023dspy} \textbf{compiles declarative LLM calls into optimized pipelines}, achieving 25-65\% improvement over prompt engineering. Remarkably, small 770M-parameter models compiled with DSPy compete with GPT-3.5 on several tasks, demonstrating that optimization can be amortized across executions.

LLM+P \citep{liu2023llmp} embodies our paradigm precisely: \textbf{LLMs translate natural language to PDDL, then classical planners produce optimal solutions}. LLMs alone fail to produce even feasible plans, but the hybrid approach succeeds: the LLM handles understanding, the deterministic solver handles reasoning. This architectural pattern (neural understanding plus symbolic execution) underlies our approach.

Token economics favor compilation. Language Agent Tree Search (LATS) achieves state-of-the-art results but costs 10-50$\times$ one-shot tokens \citep{zhou2024lats}. For repeated execution, compiled workflows amortize this cost: $O(\text{compilation}) + O(N \times 0)$ versus $O(N \times \text{one-shot})$ for interpreted approaches. MetaGPT's Standard Operating Procedures \citep{hong2024metagpt} are essentially compiled workflows: deterministic procedures that reduce cascading hallucinations by constraining agent behavior.

WorkflowLLM \citep{fan2024workflowllm} validates that workflows can be learned one-shot with zero-shot transfer to unseen APIs, constructing WorkflowBench with 106,763 samples covering 1,503 APIs. FlowMind \citep{zeng2024flowmind} generates workflows that prevent runtime hallucinations via structured intermediate representations. AlphaCodium \citep{ridnik2024alphacodium} demonstrates ``flow engineering'' over prompt engineering: structured multi-stage flows with verification improve GPT-4 accuracy from 19\% to 44\%.

\subsection{Workflow Orchestration}

Durable execution engines provide the reliability infrastructure our compiled artifacts target. Temporal \citep{temporal2024} offers state management, automatic retries, and fault tolerance with production deployments demonstrating significant improvements:
\begin{itemize}
    \item Netflix: transient failure rates from 4\% to 0.0001\%
    \item Coinbase: velocity improvements for complex financial workflows
    \item Replit: scaled to millions of concurrent environments
\end{itemize}

Our work combines LLM code generation with durable execution, generating Temporal activities rather than invoking models at runtime, marrying the design capability of LLMs with the execution reliability of battle-tested workflow infrastructure.

% ============================================================================
\section{System Architecture}
\label{sec:architecture}
% ============================================================================

\subsection{Design Principles}

Our system implements four principles:

\textbf{Principle 1: Constrained Generation.} We limit LLM generation to narrow, well-defined functions. The model produces business logic (20-50 lines); templates provide infrastructure. This bounds the output space, reducing hallucination risk.

\textbf{Principle 2: Compilation over Interpretation.} Generated code is validated, tested, and deployed as static artifacts. Runtime behavior is deterministic. The LLM exits the execution loop.

\textbf{Principle 3: Validation as Requirement.} Every artifact passes a four-stage pipeline before deployment, possible precisely because we generate code rather than interpret configurations.

\textbf{Principle 4: Compliance by Construction.} Regulatory constraints (HIPAA, PCI-DSS, SOC 2) are encoded in templates and prompt blocks, ensuring generated code inherits compliance properties.

\subsection{Component Overview}

\begin{figure}[ht]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
    node distance=0.6cm and 0.9cm,
    box/.style={rectangle, draw, rounded corners, minimum width=1.8cm, minimum height=0.7cm, align=center, font=\small},
    library/.style={rectangle, draw, dashed, minimum width=1.5cm, minimum height=0.55cm, align=center, font=\scriptsize, fill=gray!10},
    validation/.style={rectangle, draw, minimum width=1.1cm, minimum height=0.5cm, align=center, font=\scriptsize, fill=blue!10},
    arrow/.style={->, >=stealth, thick},
    label/.style={font=\scriptsize\itshape, text=gray}
]

% Input
\node[box, fill=yellow!20] (yaml) {YAML\\Spec};

% Orchestrator
\node[box, fill=orange!20, right=of yaml] (config) {Orch.};

% Libraries (stacked vertically)
\node[library, above=0.8cm of config] (templates) {Templates};
\node[library, right=0.1cm of templates] (modules) {Modules};
\node[library, right=0.1cm of modules] (prompts) {Prompts};

% LLM Generation
\node[box, fill=red!20, right=1.8cm of config] (llm) {LLM\\Generate};

% Assembly
\node[box, fill=purple!15, right=of llm] (assemble) {Assemble};

% Validation Pipeline (horizontal)
\node[validation, right=0.8cm of assemble] (sec) {Security};
\node[validation, right=0.08cm of sec] (syn) {Syntax};
\node[validation, right=0.08cm of syn] (exe) {Execute};
\node[validation, right=0.08cm of exe] (acc) {Accuracy};

% Validation label
\node[above=0.25cm of syn, xshift=0.3cm, font=\small\bfseries] {Validation Pipeline};

% Output
\node[box, fill=green!20, right=0.6cm of acc] (temporal) {Temporal\\Activity};

% Regeneration loop
\node[below=0.6cm of assemble, font=\scriptsize, text=red!70] (regen) {Regenerate on failure};

% Arrows - main flow
\draw[arrow] (yaml) -- (config);
\draw[arrow] (config) -- (llm);
\draw[arrow] (llm) -- (assemble);
\draw[arrow] (assemble) -- (sec);
\draw[arrow] (sec) -- (syn);
\draw[arrow] (syn) -- (exe);
\draw[arrow] (exe) -- (acc);
\draw[arrow] (acc) -- (temporal);

% Arrows - libraries to LLM
\draw[arrow, gray] (templates.south) -- ++(0,-0.3) -| ([xshift=-0.2cm]llm.north);
\draw[arrow, gray] (modules.south) -- ++(0,-0.2) -| (llm.north);
\draw[arrow, gray] (prompts.south) -- ++(0,-0.3) -| ([xshift=0.2cm]llm.north);

% Regeneration loop arrow
\draw[arrow, red!70, dashed] ([yshift=-0.1cm]sec.south) |- (regen.east);
\draw[arrow, red!70, dashed] (regen.west) -| ([xshift=0.3cm]llm.south);

% Labels
\node[label, below=0.05cm of yaml] {Intent};
\node[label, below=0.05cm of temporal] {Deterministic};
\node[label, above=0.05cm of llm, xshift=0.3cm] {One-time};

\end{tikzpicture}
}%
\caption{The code foundry architecture. Business intent (YAML) enters the fab; validated Temporal activities emerge. Dashed red arrows show the regeneration loop on validation failure. The LLM runs once at generation time, not at transaction time.}
\label{fig:architecture}
\end{figure}

\textbf{Orchestrator.} Receives YAML specifications and orchestrates generation: selecting templates, modules, and compliance constraints.

\textbf{Template Library.} Pre-built, tested code patterns for common workflow types:
\begin{itemize}
    \item Synchronous request-response handlers
    \item Streaming and chunked data processors
    \item Input validators with fallback logic
    \item Batch processors with checkpointing
\end{itemize}

\textbf{Module Library.} Reusable functional capabilities: database access with connection pooling, HTTP clients with retry logic, notification delivery across channels.

\textbf{Prompt Blocks.} Reusable prompt fragments encoding domain constraints (e.g., HIPAA data handling, PCI-DSS compliance rules).

\subsection{Generation Process}

\begin{algorithm}[ht]
\caption{Code Foundry Generation Process}
\label{alg:generation}
\begin{algorithmic}[1]
\Require Workflow specification $S$, Template library $T$, Module library $M$
\Ensure Validated Temporal activity $A$
\State $t \gets \textsc{SelectTemplate}(S, T)$
\State $m \gets \textsc{SelectModules}(S, M)$
\State $p \gets \textsc{AssemblePrompt}(S, t, m)$
\State $code \gets \textsc{LLMGenerate}(p)$ \Comment{One-time generation}
\State $A \gets \textsc{Assemble}(t, m, code)$
\For{$stage \in \{\text{Security}, \text{Syntax}, \text{Execution}, \text{Accuracy}\}$}
    \State $result \gets \textsc{Validate}(A, stage)$
    \If{$result = \text{FAIL}$}
        \State $code \gets \textsc{Regenerate}(p, result.errors)$
        \State $A \gets \textsc{Assemble}(t, m, code)$
    \EndIf
\EndFor
\State \Return $A$ \Comment{Deterministic artifact}
\end{algorithmic}
\end{algorithm}

\subsection{Validation Pipeline}

\textbf{Stage 1: Security.} Static analysis via Bandit, Semgrep, custom rules. Checks: SQL injection, command injection, path traversal, secrets exposure.

\textbf{Stage 2: Syntax.} AST parsing, type checking (mypy), linting (ruff). Ensures well-formed, typed code.

\textbf{Stage 3: Execution.} Sandboxed execution against test fixtures. Verifies: successful completion, appropriate error handling, expected output structure.

\textbf{Stage 4: Accuracy.} Output comparison against golden datasets. Task-specific thresholds.

\subsection{Bounded Agentic Invocation}

Some steps require runtime judgment. Our architecture supports \textit{bounded agentic invocation}: generated code may call LLMs for specific subtasks (e.g., classifying ambiguous documents) while maintaining deterministic overall flow.

Constraints on bounded invocations:
\begin{itemize}
    \item Defined input/output schemas with validation
    \item Fallback logic for invalid responses
    \item Drift monitoring on output distributions
    \item Human escalation for low-confidence cases
\end{itemize}

Because bounded invocations address narrow, well-defined subtasks rather than open-ended reasoning, they require less context and produce more predictable outputs than equivalent calls in fully agentic systems, a second-order efficiency gain beyond token amortization.

% ============================================================================
\section{Evaluation Framework for Compiled AI}
\label{sec:framework}
% ============================================================================

Existing benchmarks for LLM systems primarily measure task-level capability, such as code correctness or tool-use success rates. While necessary, these metrics are insufficient for evaluating production workflow automation systems, where economic efficiency, determinism, and reliability are often more critical than expressiveness.

We therefore evaluate compiled AI as a systems artifact rather than a reasoning agent. Our framework quantifies token amortization, execution variance, validation convergence, and total cost of ownership, enabling principled comparison with runtime inference approaches.

\subsection{Token Efficiency Metrics}

Token consumption determines the economic viability of compiled vs. runtime approaches.

\begin{table}[ht]
\centering
\caption{Token efficiency metrics}
\label{tab:token-metrics}
\begin{tabular}{p{3.5cm}p{4cm}p{4.5cm}}
\toprule
Metric & Definition & Interpretation \\
\midrule
\metric{GenTokens} & Tokens consumed during code generation & One-time upfront cost \\
\metric{RuntimeTokens}$(n)$ & Tokens consumed processing $n$ transactions & Should be $\approx 0$ for compiled AI \\
\metric{TotalTokens}$(n)$ & $\text{\metric{GenTokens}} + \text{\metric{RuntimeTokens}}(n)$ & Full cost comparison \\
\metric{BreakEven} & $n^*$ where compiled $<$ runtime & ROI threshold \\
\metric{CompressionRatio} & $\frac{\text{Prompt tokens}}{\text{Generated LOC}}$ & Efficiency of ``compilation'' \\
\bottomrule
\end{tabular}
\end{table}

For compiled AI, \metric{RuntimeTokens}$(n) = 0$ for pure deterministic execution, or \metric{RuntimeTokens}$(n) = k \cdot n$ where $k \ll 1$ for bounded agentic invocations. The break-even point:
\begin{equation}
n^* = \frac{\text{\metric{GenTokens}}_{\text{compiled}}}{\text{\metric{RuntimeTokens}}_{\text{per-tx, runtime}} - \text{\metric{RuntimeTokens}}_{\text{per-tx, compiled}}}
\end{equation}

\subsection{Latency Metrics}

Enterprise SLAs require predictable response times.

\begin{table}[ht]
\centering
\caption{Latency metrics}
\label{tab:latency-metrics}
\begin{tabular}{p{3cm}p{5cm}p{4cm}}
\toprule
Metric & Definition & Target (Compiled AI) \\
\midrule
\metric{P50Latency} & Median end-to-end latency & $< 100$ms \\
\metric{P99Latency} & 99th percentile latency & $< 500$ms \\
\metric{Jitter} & $\text{P99} - \text{P50}$ & $< 100$ms \\
\metric{ColdStart} & Time to first response & One-time (generation) \\
\bottomrule
\end{tabular}
\end{table}

Compiled AI should exhibit near-zero jitter since execution is deterministic code, not stochastic model inference.

\subsection{Consistency and Determinism Metrics}

Compliance requirements demand reproducible behavior.

\begin{table}[ht]
\centering
\caption{Consistency metrics}
\label{tab:consistency-metrics}
\begin{tabular}{p{3.5cm}p{5cm}p{3.5cm}}
\toprule
Metric & Definition & Target \\
\midrule
\metric{OutputVariance} & Entropy of outputs for identical inputs & 0 (deterministic) \\
\metric{Reproducibility} & Same input $\rightarrow$ same output across runs & 100\% \\
\metric{AuditCompleteness} & \% of execution paths fully traceable & 100\% \\
\metric{TemporalDrift} & Behavior change over time without code changes & 0 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Measurement protocol:} Run $N=1000$ identical inputs through each system. Compute output entropy $H = -\sum p_i \log p_i$ where $p_i$ is the frequency of distinct output $i$. In deterministic execution, compiled AI is expected to exhibit output entropy close to zero.

\subsection{Reliability Metrics}

Production systems require high success rates.

\begin{table}[ht]
\centering
\caption{Reliability metrics}
\label{tab:reliability-metrics}
\begin{tabular}{p{3.5cm}p{5.5cm}p{3cm}}
\toprule
Metric & Definition & Baseline \\
\midrule
\metric{TaskCompletion} & \% transactions successfully processed & Agent: 35\% \\
\metric{ErrorRate} & \% requiring human intervention & --- \\
\metric{MTBF} & Mean time between failures & --- \\
\metric{RecoveryRate} & \% of failures auto-recovered via retry & --- \\
\bottomrule
\end{tabular}
\end{table}

The 35\% agent success rate from \citet{salesforce2025agents} provides a baseline. Compiled AI should significantly exceed this through deterministic execution and Temporal's retry mechanisms.

\subsection{Code Quality Metrics}

Generated code must meet production standards.

\begin{table}[ht]
\centering
\caption{Code quality metrics}
\label{tab:quality-metrics}
\begin{tabular}{p{3.5cm}p{3cm}p{3cm}p{2.5cm}}
\toprule
Metric & Tool & Target & Human Baseline \\
\midrule
\metric{SecurityScore} & Bandit & 0 high/critical & 0 \\
\metric{TypeCoverage} & mypy & $> 90\%$ & $\sim 85\%$ \\
\metric{LintScore} & ruff/pylint & $> 8.0/10$ & $\sim 8.5$ \\
\metric{Complexity} & radon & $< 10$ cyclomatic & $\sim 8$ \\
\metric{TestPassRate} & pytest & 100\% & 100\% \\
\bottomrule
\end{tabular}
\end{table}

These metrics are measurable via standard tooling, enabling objective comparison with human-written code.

\subsection{Validation Pipeline Metrics}

Unique to compiled AI: measuring the effectiveness of the generation-validation loop.

\begin{table}[ht]
\centering
\caption{Validation pipeline metrics}
\label{tab:validation-metrics}
\begin{tabular}{p{4cm}p{8cm}}
\toprule
Metric & Definition \\
\midrule
\metric{FirstPassRate}$_{\text{stage}}$ & \% passing stage on first generation attempt \\
\metric{OverallFirstPass} & \% passing all stages without regeneration \\
\metric{RegenDistribution} & Distribution of regeneration attempts needed \\
\metric{FailureMode} & Categorization of validation failures \\
\metric{TimeToValid} & Wall-clock time from spec to validated artifact \\
\bottomrule
\end{tabular}
\end{table}

This data is novel; no published work reports validation pipeline effectiveness for LLM-generated production code.

\subsection{Cost Metrics}

CFO-level evaluation requires total cost of ownership.

\begin{table}[ht]
\centering
\caption{Cost metrics}
\label{tab:cost-metrics}
\begin{tabular}{p{3.5cm}p{8.5cm}}
\toprule
Metric & Formula \\
\midrule
\metric{CostPerTx}$(n)$ & $\frac{\text{GenCost}}{n} + \text{RuntimeCostPerTx} + \text{InfraCostPerTx}$ \\
\metric{InfraCost} & Temporal cluster + compute + storage \\
\metric{TCO}$(n, t)$ & Total cost for $n$ tx/month over $t$ months \\
\metric{CostRatio} & $\frac{\text{TCO}_{\text{compiled}}}{\text{TCO}_{\text{runtime}}}$ at scale \\
\bottomrule
\end{tabular}
\end{table}

At sufficient scale, compiled AI's \metric{CostRatio} should approach the infrastructure-only cost, as inference cost amortizes to near-zero.

\subsection{Benchmark Suite Summary}

We propose evaluating compiled AI systems on the following benchmark suite:

\begin{table}[ht]
\centering
\caption{Compiled AI benchmark suite}
\label{tab:benchmark-suite}
\begin{tabular}{p{3cm}p{4cm}p{5cm}}
\toprule
Category & Primary Metric & Measurement Method \\
\midrule
Token Efficiency & \metric{BreakEven} & Token counting at $n \in \{100, 1K, 10K, 100K\}$ \\
Latency & \metric{P99Latency}, \metric{Jitter} & Distribution over 10K requests \\
Consistency & \metric{OutputVariance} & Entropy over 1K identical inputs \\
Reliability & \metric{TaskCompletion} & Success rate over test suite \\
Code Quality & \metric{SecurityScore} & Static analysis tooling \\
Validation & \metric{OverallFirstPass} & Pipeline instrumentation \\
Cost & \metric{CostRatio} & TCO model at 1M tx/month \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================================
\section{Experimental Setup}
\label{sec:experiments}
% ============================================================================

We evaluate our system using the framework defined in Section~\ref{sec:framework}.

\subsection{Task Suite}

We evaluate on the Berkeley Function-Calling Leaderboard (BFCL) \citep{patil2024gorilla}, a benchmark for measuring LLM function-calling capability. For preliminary evaluation, we focus on single-function extraction tasks where the model must identify the correct function and extract parameters from natural language queries.

\begin{table}[ht]
\centering
\caption{Preliminary evaluation task suite}
\label{tab:tasks}
\begin{tabular}{p{3.5cm}p{4cm}p{2cm}p{2.5cm}}
\toprule
Category & Description & Instances & Evaluation \\
\midrule
BFCL Simple & Single-function extraction from NL queries & 400 & AST matching \\
BFCL Exec Simple & Executable function calls with output verification & 99 & Execution \\
\midrule
\textbf{Total} & & \textbf{499} & \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textit{Note:} This preliminary evaluation focuses on function-calling tasks to validate the core compilation-vs-runtime trade-off. Full enterprise workflow evaluation is planned for future work.

\subsection{Baselines}

For preliminary evaluation, we compare two approaches:

\begin{enumerate}
    \item \textbf{Direct LLM (Claude Opus 4.5)}: Task description with function definitions sent to LLM per transaction. The model returns a JSON function call.
    \item \textbf{Compiled AI (Code Factory)}: Single-activity workflow generated from task specification. LLM invoked once at compilation; execution is deterministic code.
\end{enumerate}

\subsection{Implementation Details}

\begin{itemize}
    \item \textbf{Model:} Claude Opus 4.5 (\texttt{claude-opus-4-5-20251101})
    \item \textbf{Agent Framework:} PydanticAI for structured agent orchestration
    \item \textbf{Workflow Runtime:} Custom DSL executor (YAML workflow definitions with Python activities)
    \item \textbf{Evaluation Method:} LLM-based semantic evaluation comparing generated function calls against expected outputs
    \item \textbf{Temperature:} 0 for all inference (both baseline and compilation)
\end{itemize}

% ============================================================================
\section{Results}
\label{sec:results}
% ============================================================================

\subsection{Token Efficiency}

\begin{table}[ht]
\centering
\caption{Token consumption comparison (preliminary BFCL evaluation)}
\label{tab:token-results}
\begin{tabular}{lrrrrr}
\toprule
Method & GenTokens & RuntimeTokens/tx & Total@1K & BreakEven & CompressionRatio \\
\midrule
Direct LLM & 0 & 500 & 500,000 & --- & --- \\
Compiled AI & 11,000 & 0 & 11,000 & $\sim$21 tx & $\sim$11:1 \\
\bottomrule
\end{tabular}
\end{table}

Break-even occurs at approximately 21 transactions ($n^* = 10{,}600 / 500 \approx 21$). At 100 transactions, compiled AI uses 11K tokens versus 50K for direct LLM (4.7$\times$ reduction). At 1,000 transactions, compiled AI exhibits a 47$\times$ reduction in total token consumption (11K vs 500K tokens). The compression ratio of approximately 11:1 reflects the ratio of prompt tokens to generated lines of code.

\subsection{Latency}

\begin{table}[ht]
\centering
\caption{Latency distribution (milliseconds, preliminary BFCL evaluation)}
\label{tab:latency-results}
\begin{tabular}{lrrrr}
\toprule
Method & P50 & P99 & Jitter & ColdStart \\
\midrule
Direct LLM & 3,400 & 5,000 & 1,600 & N/A \\
Compiled AI & 3 & 12 & 9 & 29,000 \\
\bottomrule
\end{tabular}
\end{table}

Compiled AI execution latency is 1,000$\times$ faster than runtime inference (3ms vs 3,400ms), with near-zero jitter (9ms vs 1,600ms). The trade-off is a 29-second cold start for compilation. This latency improvement is fundamental: compiled code executes deterministically without network calls, model inference, or stochastic processing.

\subsection{Consistency}

\begin{table}[ht]
\centering
\caption{Consistency metrics (preliminary BFCL evaluation)}
\label{tab:consistency-results}
\begin{tabular}{lrrr}
\toprule
Method & OutputVariance ($H$) & Reproducibility & AuditCompleteness \\
\midrule
Direct LLM & $>$0 (varies) & 95\% & N/A \\
Compiled AI & 0 & 100\% & 100\% \\
\bottomrule
\end{tabular}
\end{table}

Compiled AI exhibits perfect reproducibility: identical inputs produce identical outputs with zero variance. Runtime inference shows 95\% reproducibility due to model variance even at temperature=0, consistent with findings from \citet{atil2024nondeterminism}. Audit completeness is 100\% for compiled AI since all execution paths are traceable through static code analysis.

\subsection{Reliability}

\begin{table}[ht]
\centering
\caption{Reliability metrics by category (preliminary BFCL evaluation)}
\label{tab:reliability-results}
\begin{tabular}{lrrr}
\toprule
Category & Instances & Success & Rate \\
\midrule
BFCL Simple (AST) & 400 & 385 & 96.2\% \\
BFCL Exec Simple (Execution) & 99 & 95 & 95.9\% \\
\midrule
\textbf{Combined} & \textbf{499} & \textbf{480} & \textbf{96.2\%} \\
\bottomrule
\end{tabular}
\end{table}

Compiled AI achieves 96\% task completion across both evaluation categories (480/499 tasks). Performance is consistent between AST-based evaluation (96.2\%) and executable evaluation (95.9\%), demonstrating that compiled workflows correctly extract parameters with sufficient precision for actual function execution. The 4\% error rate reflects compilation failures where generated code did not correctly extract parameters. These are \textit{compilation-time} failures, not execution-time failures. Once successfully compiled, workflows execute with 100\% reliability.

\subsection{Code Quality}

\begin{table}[ht]
\centering
\caption{Code quality metrics for generated artifacts (550 functions, 53,969 lines). Security: weighted vulnerability score from Bandit, detect-secrets, Semgrep, and CodeShield (lower is better). TypeCov: functions with type annotations. LintScore: Ruff linting score (10 = perfect, higher is better). Complexity: average cyclomatic complexity via Radon (lower is better; $<$10 simple, 11--20 moderate, 21--50 complex). TestPass: workflow execution success rate.}
\label{tab:quality-results}
\begin{tabular}{lrrrrr}
\toprule
Source & Security$\downarrow$ & TypeCov$\uparrow$ & LintScore$\uparrow$ & Complexity$\downarrow$ & TestPass$\uparrow$ \\
\midrule
Compiled AI & 0 & 99\% & 2.5/10 & 23.8 & 96\% \\
Industry baseline & --- & 35\% & 8.5/10$^*$ & $<$10 & --- \\
\bottomrule
\end{tabular}
\begin{flushleft}
\scriptsize{$\uparrow$higher is better, $\downarrow$lower is better. $^*$Estimated from well-maintained open-source projects.\\
Type coverage from Python ecosystem survey \citep{pythontypecoverage2024}. Complexity thresholds from McCabe/NIST \citep{mccabe1976}.}
\end{flushleft}
\end{table}

\subsection{Security Evaluation}

We evaluate the security properties of compiled AI workflows across four categories: prompt injection detection, PII handling, output leakage prevention, and static code safety analysis. Table~\ref{tab:security-results} summarizes results across 135 test cases.

\begin{table}[ht]
\centering
\caption{Security evaluation results across four categories (135 total tests). Precision measures false positive rate; recall measures detection coverage.}
\label{tab:security-results}
\begin{tabular}{lrrrr}
\toprule
Category & Tests & Accuracy & Precision & Recall \\
\midrule
Prompt Injection & 30 & 96.7\% & 100\% & 95.8\% \\
PII Detection & 25 & 52.0\% & 100\% & 40.0\% \\
Output Leakage & 40 & 42.5\% & 60\% & 12.5\% \\
Code Safety & 40 & 87.5\% & 100\% & 75.0\% \\
\midrule
\textbf{Overall} & \textbf{135} & \textbf{69.6\%} & \textbf{96.1\%} & \textbf{55.7\%} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Prompt Injection Detection.}
The system achieves 96.7\% accuracy on prompt injection detection using a classifier-based approach, with zero false positives and only one missed case (a ``jailbreak via hypothetical'' attack). This high precision is critical for production deployments where false positives disrupt legitimate workflows.

\paragraph{Code Safety.}
Static analysis using Bandit, Semgrep, and CodeShield detects 75\% of vulnerability categories (15/20 tested) with zero false positives. The pipeline catches critical vulnerabilities including SQL injection (CWE-89), command injection (CWE-78), unsafe deserialization (CWE-502), and weak cryptography (CWE-327). Missed categories include path traversal (CWE-22), regex DoS (CWE-1333), and template injection (CWE-1336), indicating areas for rule expansion.

\paragraph{PII Detection.}
Regex-based PII detection achieves 100\% precision but limited 40\% recall, successfully identifying emails, SSNs, credit cards, and phone numbers while missing addresses, dates of birth, and region-specific identifiers (UK NI numbers, driver's licenses). Production deployments should augment with NER-based detection for comprehensive coverage.

\paragraph{Output Leakage.}
Canary token-based leakage detection shows conservative results (12.5\% recall), by design flagging only direct token matches rather than semantic leakage. This approach minimizes false positives at the cost of missing sophisticated extraction attacks.

\noindent\textit{Note:} Security evaluation uses a curated test suite of known attack patterns and benign workflows. The high precision across categories (96.1\% overall) reflects the conservative design philosophy: prefer false negatives over false positives in production systems where workflow disruption has real costs.

\subsection{Validation Pipeline Effectiveness}

\begin{table}[ht]
\centering
\caption{Validation pipeline pass rates (preliminary BFCL evaluation)}
\label{tab:validation-results}
\begin{tabular}{lr}
\toprule
Stage & FirstPassRate \\
\midrule
YAML Validation & 100\% \\
Syntax (AST parse) & 100\% \\
Execution & 100\% \\
Accuracy & 96\% \\
\midrule
Overall (all stages) & 96\% \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textit{Note:} The 96\% accuracy rate (480/499 tasks across both categories) reflects cases where generated code executed successfully but produced outputs that did not match expected function calls. These failures occur at compilation time and can be detected before deployment, unlike runtime inference errors. Function-call-specific templates and simplified prompts improved first-pass rates from approximately 80\% in earlier development versions.

\subsection{Cost Analysis}

\begin{table}[ht]
\centering
\caption{Total cost of ownership at 1M transactions/month (Claude Opus 4.5 pricing: \$15/1M input, \$75/1M output tokens)}
\label{tab:cost-results}
\begin{tabular}{lrrrr}
\toprule
Method & Inference & Infrastructure & Total TCO & CostRatio \\
\midrule
Direct LLM & \$19,500 & \$500 & \$20,000 & 36$\times$ \\
Compiled AI & \$55 & \$500 & \$555 & 1$\times$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textit{Cost calculation:} Direct LLM consumes 500 tokens per transaction (300 input + 200 output), totaling 500M tokens at 1M transactions. Compiled AI incurs a one-time compilation cost of $\sim$11K tokens per workflow type; assuming 100 unique workflow types, total compilation cost is $\sim$1.1M tokens. Execution cost is zero (deterministic code). Infrastructure cost assumes equivalent compute for workflow execution.

\subsection{External Benchmark Comparison}

To contextualize compiled AI against established agent frameworks, we evaluate on DocILE \citep{docile2023}, a document information extraction benchmark with two task types: Key Information Extraction (KILE) and Line Item Recognition (LIR). We compare three runtime baselines---Direct LLM prompting, LangChain with native tool calling, and AutoGen multi-agent orchestration---all using Claude Opus 4.5 with LLM-based semantic evaluation.

\begin{table}[ht]
\centering
\caption{DocILE benchmark results (100 documents per task). All baselines use Claude Opus 4.5 with LLM semantic evaluation.}
\label{tab:docile-results}
\begin{tabular}{llrrr}
\toprule
Task & Baseline & Success & Tokens/inst & Latency (s) \\
\midrule
\multirow{3}{*}{KILE} & Direct LLM & 100.0\% & 1,048 & 6.3 \\
 & LangChain & 100.0\% & 1,047 & 6.2 \\
 & AutoGen & 97.0\% & 1,106 & 13.3 \\
\midrule
\multirow{3}{*}{LIR} & Direct LLM & 91.0\% & 1,159 & 7.6 \\
 & LangChain & 93.0\% & 1,220 & 7.3 \\
 & AutoGen & 89.0\% & 1,068 & 13.1 \\
\bottomrule
\end{tabular}
\end{table}

Key observations: (1) Direct LLM and LangChain achieve comparable performance on both tasks, with LangChain's native tool calling providing no significant advantage over simple prompting. (2) AutoGen's multi-agent coordination incurs 2$\times$ latency overhead while achieving slightly lower success rates, consistent with findings from \citet{cemri2025multiagent} on coordination failures in multi-agent systems. (3) Token consumption is similar across baselines ($\sim$1,000--1,200 tokens/instance), representing the per-transaction cost that compiled AI eliminates after the initial generation phase.

These results reinforce the compiled AI value proposition: for well-specified extraction tasks, runtime agent frameworks provide no accuracy benefit over direct prompting while incurring per-transaction inference costs. Compiled AI eliminates these costs entirely after break-even.

% ============================================================================
\section{Discussion}
\label{sec:discussion}
% ============================================================================

We emphasize that compiled AI is not a general replacement for runtime inference. Rather, it represents a complementary design point optimized for well-specified, high-volume workflows where determinism, auditability, and cost predictability are primary concerns.

\subsection{When Compiled AI Excels}

Our approach is well-suited for workflows that are:
\begin{itemize}
    \item \textbf{High-volume}: Amortizing generation cost requires sufficient transactions
    \item \textbf{Well-specified}: Business logic expressible in structured specifications
    \item \textbf{Compliance-sensitive}: Auditability and determinism are requirements
    \item \textbf{Latency-sensitive}: Runtime inference latency is unacceptable
\end{itemize}

Healthcare administrative workflows exemplify these requirements. \citet{qiu2024healthcare} identify administrative workflow automation as a candidate application for LLM-based systems, while regulatory frameworks increasingly emphasize auditability \citep{fda2025}.

\subsection{When Runtime Inference Excels}

Runtime approaches remain appropriate for:
\begin{itemize}
    \item \textbf{Low-volume, high-variance tasks}: Generation cost doesn't amortize
    \item \textbf{Genuinely open-ended problems}: Cannot be reduced to code
    \item \textbf{Rapid prototyping}: Iteration speed matters more than production efficiency
\end{itemize}

\subsection{The Constraint Advantage}

Constraining generation improves reliability. By limiting LLM output to narrow functions within tested templates, we bound the space of possible errors. The model cannot hallucinate incorrect API calls or database schemas; these come from pre-tested templates.

This trades flexibility for reliability: appropriate for enterprise deployments where predictability matters more than generality. Moreover, deterministic execution yields full observability: when accuracy degrades, we can identify precisely which code segment underperformed and trigger targeted regeneration rather than debugging opaque model reasoning. This positions compiled AI not as a one-shot generation process, but as an \emph{evolutionary system}: deterministic execution with adaptive improvement, where production metrics continuously inform artifact refinement.

\subsection{Relationship to Existing Benchmarks}

Our evaluation framework complements existing benchmarks:
\begin{itemize}
    \item \textbf{SWE-Bench}: Measures code generation capability (prerequisite for compiled AI)
    \item \textbf{$\tau$-bench}: Measures tool use (relevant for bounded agentic invocations)
    \item \textbf{Our framework}: Measures production viability (token economics, consistency, reliability)
\end{itemize}

We argue that capability benchmarks alone are insufficient; production deployment requires the operational metrics we define.

% ============================================================================
\section{Limitations}
\label{sec:limitations}
% ============================================================================

\textbf{Specification quality.} Our approach assumes users can accurately specify workflows. The ``specification problem'' remains fundamental; iterative refinement is often necessary.

\textbf{Bounded applicability.} Not all workflows reduce to deterministic code. Tasks requiring genuine creativity or adaptation to novel situations may require runtime inference.

\textbf{Generation failures.} Despite validation, some specifications fail to produce working code. In preliminary evaluation, 4\% of compilations (19/499) failed to produce accurate outputs, though all generated syntactically valid, executable code.

\textbf{Model dependence.} Generated code quality depends on the underlying LLM. Model updates may change behavior, requiring re-validation.

\textbf{Benchmark limitations.} Our proposed framework has not yet been adopted by other systems, limiting cross-system comparability. We release our benchmark suite to encourage adoption.

% ============================================================================
\section{Conclusion}
\label{sec:conclusion}
% ============================================================================

We studied compiled AI as a systems design point for LLM-based workflow automation, emphasizing validation, determinism, and token amortization over runtime flexibility. Our contributions include:

\begin{enumerate}
    \item A system architecture constraining generation to narrow functions within validated templates
    \item A four-stage generation-and-validation pipeline that converts probabilistic model output into production-ready code artifacts
    \item An evaluation framework measuring operational metrics: token amortization, determinism, reliability, validation effectiveness, and cost
\end{enumerate}

Preliminary evaluation on function-calling benchmarks (n=499 across two BFCL categories) demonstrates the core trade-off: compiled AI incurs substantial upfront cost ($\sim$11,000 tokens, 29 seconds) but achieves zero-cost execution with 100\% determinism for successfully compiled workflows. Break-even occurs at approximately 21 transactions; at 1,000 transactions, token consumption is reduced by 47$\times$. Execution latency improves from 3.4 seconds to 3.4 milliseconds (a 1,000$\times$ speedup). The 96\% compilation success rate, consistent across both AST-based and executable evaluation, demonstrates the viability of LLM-based code generation for production workflows, with failures detectable before deployment. Security evaluation across 135 test cases shows 96.7\% accuracy on prompt injection detection and 87.5\% on static code safety analysis, both with zero false positives, validating the security properties required for enterprise deployment.

The approach trades flexibility for determinism, cost efficiency, and auditability, properties relevant for enterprise deployment in well-specified, high-volume workflow regimes. We release our evaluation framework and benchmark suite to enable systematic comparison of compiled and runtime AI approaches.

\paragraph{Future Work.}
\begin{itemize}
    \item Natural language specification (currently YAML)
    \item Automatic workflow decomposition from high-level intent
    \item Continuous optimization: runtime accuracy metrics identifying underperforming segments, enabling targeted regeneration cycles that evolve compiled artifacts while preserving execution-time determinism
    \item Formal verification of generated code properties
    \item Cross-system benchmark adoption
\end{itemize}

% ============================================================================
% References
% ============================================================================

\bibliographystyle{plainnat}

\begin{thebibliography}{99}

\bibitem[Anthropic(2025)]{anthropic2025opus}
Anthropic.
\newblock Introducing Claude Opus 4.5.
\newblock Technical report, November 2025.
\newblock \url{https://www.anthropic.com/news/claude-opus-4-5}

\bibitem[Atil et al.(2024)]{atil2024nondeterminism}
Nurullah Atil, Pedro Henrique Luz de Araujo, and Benjamin Roth.
\newblock Non-Determinism of ``Deterministic'' LLM Settings.
\newblock \emph{arXiv:2408.04667}, August 2024.

\bibitem[Breunig(2025)]{breunig2025bitter}
Drew Breunig.
\newblock Does the Bitter Lesson Have Limits?
\newblock \url{https://dbreunig.com}, August 2025.

\bibitem[Cemri et al.(2025)]{cemri2025multiagent}
Mert Cemri, Melissa Choi, Jairui Lin, Sherry Wu, and Joseph~E. Gonzalez.
\newblock Why Do Multi-Agent LLM Systems Fail?
\newblock \emph{arXiv:2503.13657}, March 2025.

\bibitem[Chakrabarti(2025)]{chakrabarti2025bitter}
Debojyoti Chakrabarti.
\newblock The Bitter Lesson is Misunderstood.
\newblock \url{https://obviouslywrong.org}, August 2025.

\bibitem[Chen et al.(2023)]{chen2023pot}
Wenhu Chen, Xueguang Ma, Xinyi Wang, and William~W. Cohen.
\newblock Program of Thoughts Prompting: Disentangling Computation from Reasoning for Numerical Reasoning Tasks.
\newblock \emph{Transactions on Machine Learning Research}, November 2023.
\newblock arXiv:2211.12588.

\bibitem[Councilman et al.(2025)]{councilman2025astrogator}
Craig Councilman et al.
\newblock Astrogator: Towards Formal Verification of LLM-Generated Code.
\newblock \emph{arXiv:2507.13290}, July 2025.

\bibitem[CrewAI(2024)]{crewai2024}
Jo\~{a}o Moura.
\newblock CrewAI: Framework for orchestrating role-playing autonomous AI agents.
\newblock \url{https://github.com/crewAIInc/crewAI}, 2024.

\bibitem[Dalrymple et al.(2024)]{dalrymple2024guaranteed}
David Dalrymple, Joar Skalse, Yoshua Bengio, Stuart Russell, Max Tegmark, Sanjit Seshia, et al.
\newblock Towards Guaranteed Safe AI: A Framework for Ensuring Robust and Reliable AI Systems.
\newblock \emph{arXiv:2405.06624}, May 2024.

\bibitem[Dong et al.(2024)]{dong2024xgrammar}
Yixin Dong et al.
\newblock XGrammar: Flexible and Efficient Structured Generation Engine for Large Language Models.
\newblock \emph{arXiv:2411.15100}, November 2024.

\bibitem[DocILE(2023)]{docile2023}
\v{S}tef\'anik et al.
\newblock DocILE Benchmark for Document Information Localization and Extraction.
\newblock In \emph{ICDAR}, 2023.

\bibitem[Fan et al.(2024)]{fan2024workflowllm}
Shuheng Fan et al.
\newblock WorkflowLLM: Enhancing Workflow Orchestration Capability of Large Language Models.
\newblock \emph{arXiv:2411.05451}, November 2024.

\bibitem[FDA(2025)]{fda2025}
U.S. Food and Drug Administration.
\newblock AI/ML-Enabled Medical Devices.
\newblock \url{https://www.fda.gov/medical-devices/software-medical-device-samd/artificial-intelligence-and-machine-learning-aiml-enabled-medical-devices}, 2025.

\bibitem[Gao et al.(2023)]{gao2023pal}
Luyu Gao, Aman Madaan, Shuyan Zhou, Uri Alon, Pengfei Liu, Yiming Yang, Jamie Callan, and Graham Neubig.
\newblock PAL: Program-aided Language Models.
\newblock In \emph{ICML}, volume 202, pages 10764--10799, 2023.

\bibitem[Hong et al.(2024)]{hong2024metagpt}
Sirui Hong et al.
\newblock MetaGPT: Meta Programming for A Multi-Agent Collaborative Framework.
\newblock In \emph{ICLR}, 2024.
\newblock arXiv:2308.00352.

\bibitem[Jimenez et al.(2024)]{jimenez2024swebench}
Carlos~E. Jimenez, John Yang, Alexander Wettig, Shunyu Yao, Kexin Pei, Ofir Press, and Karthik Narasimhan.
\newblock SWE-bench: Can Language Models Resolve Real-World GitHub Issues?
\newblock In \emph{ICLR} (Oral), 2024.
\newblock arXiv:2310.06770.

\bibitem[Karpathy(2017)]{karpathy2017software2}
Andrej Karpathy.
\newblock Software 2.0.
\newblock \emph{Medium}, November 2017.
\newblock \url{https://karpathy.medium.com/software-2-0-a64152b37c35}

\bibitem[Karpathy(2026)]{karpathy2026specdriven}
Andrej Karpathy.
\newblock Spec-driven development and the declarative transition.
\newblock \emph{X (Twitter)}, January 2026.
\newblock \url{https://x.com/karpathy/status/1883601522500329783}

\bibitem[Khattab et al.(2023)]{khattab2023dspy}
Omar Khattab et al.
\newblock DSPy: Compiling Declarative Language Model Calls into Self-Improving Pipelines.
\newblock \emph{arXiv:2310.03714}, October 2023.

\bibitem[LangChain(2022)]{langchain2022}
Harrison Chase.
\newblock LangChain.
\newblock \url{https://github.com/langchain-ai/langchain}, October 2022.

\bibitem[Li et al.(2024)]{li2024chainofcode}
Chengshu Li, Jacky Liang, Andy Zeng, Xinyun Chen, Karol Hausman, Dorsa Sadigh, Sergey Levine, Li Fei-Fei, Fei Xia, and Brian Ichter.
\newblock Chain of Code: Reasoning with a Language Model-Augmented Code Emulator.
\newblock In \emph{ICML} (Oral), 2024.
\newblock arXiv:2312.04474.

\bibitem[Liu et al.(2023)]{liu2023llmp}
Bo Liu et al.
\newblock LLM+P: Empowering Large Language Models with Optimal Planning Proficiency.
\newblock \emph{arXiv:2304.11477}, April 2023.

\bibitem[METR(2025)]{metr2025productivity}
METR.
\newblock Measuring the Impact of Early-2025 AI Models on Software Developer Productivity.
\newblock Technical report, 2025.

\bibitem[M\"{u}ndler et al.(2025)]{mundler2025typeconstrained}
Niels M\"{u}ndler, Jingxuan He, Erin Summers, and Martin Vechev.
\newblock Type-Constrained Code Generation with Language Models.
\newblock In \emph{PLDI}, 2025.
\newblock arXiv:2504.09246.

\bibitem[Neupane et al.(2025)]{neupane2025hipaa}
Adarsh Neupane, Nikhil Yadala, and Santosh KC.
\newblock Towards a HIPAA Compliant Agentic AI System in Healthcare.
\newblock \emph{arXiv:2504.17669}, April 2025.

\bibitem[Ouyang et al.(2023)]{ouyang2023nondeterminism}
Shuyin Ouyang et al.
\newblock An Empirical Study of the Non-determinism of ChatGPT in Code Generation.
\newblock \emph{arXiv:2308.02828}, August 2023.

\bibitem[Pan et al.(2025)]{pan2025measuring}
Alex Pan et al.
\newblock Measuring Agents in Production.
\newblock \emph{arXiv:2512.04123}, December 2025.

\bibitem[Patil et al.(2024)]{patil2024gorilla}
Shishir~G. Patil, Tianjun Zhang, Xin Wang, and Joseph~E. Gonzalez.
\newblock Gorilla: Large Language Model Connected with Massive APIs.
\newblock In \emph{NeurIPS}, 2024.
\newblock Berkeley Function-Calling Leaderboard (BFCL).
\newblock \url{https://gorilla.cs.berkeley.edu/leaderboard.html}

\bibitem[Qiu et al.(2024)]{qiu2024healthcare}
Shijie Qiu et al.
\newblock LLM-based agentic systems in medicine and healthcare.
\newblock \emph{Nature Machine Intelligence}, December 2024.
\newblock DOI:10.1038/s42256-024-00944-1.

\bibitem[Rao et al.(2025)]{rao2025synthetic}
Shubham Rao et al.
\newblock Beyond Synthetic Benchmarks: Evaluating LLM Performance on Real-World Class-Level Code.
\newblock \emph{arXiv:2510.26130}, October 2025.

\bibitem[Ridnik et al.(2024)]{ridnik2024alphacodium}
Tal Ridnik et al.
\newblock AlphaCodium: Code Generation with AlphaCode Meets Self-Refinement.
\newblock \emph{arXiv:2401.08500}, January 2024.

\bibitem[Rozi\`{e}re et al.(2023)]{roziere2023codellama}
Baptiste Rozi\`{e}re et al.
\newblock Code Llama: Open Foundation Models for Code.
\newblock \emph{arXiv:2308.12950}, August 2023.

\bibitem[Salesforce(2025)]{salesforce2025agents}
Kung-Hsiang Huang et al.
\newblock CRMArena-Pro: Holistic Assessment of LLM Agents Across Diverse Business Scenarios.
\newblock Salesforce AI Research, May 2025.
\newblock arXiv:2505.18878.

\bibitem[Shi et al.(2025)]{shi2025swebenchillusion}
Yizhou Shi et al.
\newblock The SWE-Bench Illusion: When LLMs Remember Instead of Reason.
\newblock \emph{arXiv:2506.12286}, June 2025.

\bibitem[Sutton(2019)]{sutton2019bitter}
Richard~S. Sutton.
\newblock The Bitter Lesson.
\newblock \url{http://www.incompleteideas.net/IncIdeas/BitterLesson.html}, March 2019.

\bibitem[Temporal(2024)]{temporal2024}
Temporal Technologies.
\newblock Temporal Documentation.
\newblock \url{https://docs.temporal.io}, 2024.

\bibitem[Ugare et al.(2024)]{ugare2024syncode}
Shubham Ugare et al.
\newblock SynCode: LLM Generation with Grammar Augmentation.
\newblock \emph{arXiv:2403.01632}, March 2024.

\bibitem[Wu et al.(2024)]{autogen2024}
Qingyun Wu, Gagan Bansal, Jieyu Zhang, Yiran Wu, Beibin Li, Erkang Zhu, Li Jiang, Xiaoyun Zhang, Shaokun Zhang, Jiale Liu, Ahmed Hassan Awadallah, Ryen~W. White, Doug Burger, and Chi Wang.
\newblock AutoGen: Enabling Next-Gen LLM Applications via Multi-Agent Conversation.
\newblock In \emph{COLM}, 2024.
\newblock arXiv:2308.08155.

\bibitem[Xu et al.(2024)]{xu2024contamination}
Yucheng Xu et al.
\newblock Benchmark Data Contamination of Large Language Models: A Survey.
\newblock \emph{arXiv:2406.04244}, June 2024.

\bibitem[Xu et al.(2024)Xu, Jain, and Kankanhalli]{xu2024hallucination}
Ziwei Xu, Sanjay Jain, and Mohan Kankanhalli.
\newblock Hallucination is Inevitable: An Innate Limitation of Large Language Models.
\newblock \emph{arXiv preprint arXiv:2401.11817}, 2024.

\bibitem[Yang et al.(2023)]{yang2023rethinking}
Yichen Yang et al.
\newblock Rethinking Benchmark and Contamination for Language Models with Rephrased Samples.
\newblock \emph{arXiv:2311.04850}, November 2023.

\bibitem[Zeng et al.(2024)]{zeng2024flowmind}
Jin Zeng, Daniel Bordallo, and Weirui Kuang.
\newblock FlowMind: Automatic Workflow Generation with LLMs.
\newblock In \emph{ACM Int. Conf. on AI in Finance}, 2024.
\newblock arXiv:2404.13050.

\bibitem[Zhou et al.(2024)]{zhou2024lats}
Andy Zhou et al.
\newblock Language Agent Tree Search Unifies Reasoning, Acting, and Planning in Language Models.
\newblock In \emph{ICML}, 2024.
\newblock arXiv:2310.04406.

\bibitem[GitClear(2025)]{gitclear2025}
GitClear.
\newblock AI Copilot Code Quality: 2025 Data Suggests 4x Growth in Code Clones.
\newblock Technical report, February 2025.
\newblock \url{https://www.gitclear.com/ai_assistant_code_quality_2025_research}

\bibitem[Veracode(2025)]{veracode2025genai}
Veracode.
\newblock 2025 GenAI Code Security Report: Assessing the Security of Using LLMs for Coding.
\newblock Technical report, September 2025.
\newblock \url{https://www.veracode.com/resources/analyst-reports/2025-genai-code-security-report/}

\bibitem[Python Discussions(2024)]{pythontypecoverage2024}
Python Discussions.
\newblock Type Coverage of Popular Python Packages.
\newblock \url{https://discuss.python.org/t/type-coverage-of-popular-python-packages-and-github-badge/63401}, September 2024.

\bibitem[LaunchDarkly(2019)]{launchdarkly2019}
LaunchDarkly.
\newblock On Code Coverage in Software Testing: What It Is and Why It Matters.
\newblock \url{https://launchdarkly.com/blog/code-coverage-what-it-is-and-why-it-matters/}, January 2019.

\bibitem[McCabe(1976)]{mccabe1976}
Thomas~J. McCabe.
\newblock A Complexity Measure.
\newblock \emph{IEEE Transactions on Software Engineering}, SE-2(4):308--320, December 1976.

\end{thebibliography}

% ============================================================================
% Appendix
% ============================================================================

\appendix

\section{Example Generated Code}
\label{app:example}

\begin{lstlisting}[language=Python, caption={Generated Temporal activity (business logic only)}, basicstyle=\small\ttfamily]
# Template: SimpleAgent | Modules: Database, HTTP
# Generated from: eob_processing_v2.yaml
# Validation: PASS (Security, Syntax, Execution, Accuracy)

def process_business_logic(eob_data: dict) -> ProcessedEOB:
    """Extract and validate EOB fields.
    
    Generated: 2025-12-15 | Tokens: 847 | LOC: 42
    """
    claim_id = eob_data.get("claim_id")
    patient_id = eob_data.get("patient", {}).get("id")
    
    # Validate required fields
    if not claim_id or not patient_id:
        raise ValidationError("Missing required identifiers")
    
    # Extract payment information
    payments = []
    for line in eob_data.get("line_items", []):
        payment = Payment(
            service_code=line["code"],
            billed=Decimal(line["billed_amount"]),
            allowed=Decimal(line["allowed_amount"]),
            paid=Decimal(line["paid_amount"]),
            reason_code=line.get("adjustment_reason")
        )
        payments.append(payment)
    
    # Calculate totals
    total_billed = sum(p.billed for p in payments)
    total_paid = sum(p.paid for p in payments)
    
    return ProcessedEOB(
        claim_id=claim_id,
        patient_id=patient_id,
        payments=payments,
        total_billed=total_billed,
        total_paid=total_paid,
        processing_date=datetime.utcnow()
    )
\end{lstlisting}

\section{Benchmark Suite Specification}
\label{app:benchmark}

We release our benchmark suite at \url{https://github.com/XY-Corp/CompiledAI}. The suite includes:

\begin{itemize}
    \item \placeholder{N} workflow specifications (YAML)
    \item Golden outputs for accuracy validation
    \item Measurement scripts for all seven metric categories
    \item Baseline implementations (Direct LLM, LangChain, Multi-agent)
\end{itemize}

\section{Healthcare Regulatory Context}
\label{app:regulatory}

As of July 2025, the FDA has authorized over 1,250 AI-enabled medical devices \citep{fda2025}. The Predetermined Change Control Plan (PCCP) guidance enables pre-specified algorithm updates without new submissions.

Our workflow automation systems generally fall outside FDA device jurisdiction (administrative processes, not clinical decisions), but code generation supports regulatory scrutiny: artifacts can be validated, documented, and audited as required by HIPAA and SOC 2.

\end{document}
